# Using R as a calculator

## Challenge 2

Use R to calculate the following skill testing question: 2 + 16 x 24 -
56

```{r}
2 + 16 * 24 - 56
```


## Challenge 3

Use R to calculate the following skill testing question:
2 + 16 x 24 - 56 / (2 + 1) - 457

*Pay attention to the order of operations when thinking about this
question!*

```{r}
2 + 16 * 24 - 56 / (2 + 1) - 457
```

Note that R *always* follows the order of priorities.

**R TIP**

Try using the "up" and "down" arrows to reproduce previous commands.
These keys actually allow you to scroll through your command history.
This is a useful tool to go back and see what command you ran and if you
might have made a mistake in it. This is always a useful tool to quickly
alter previous commands you ran and to re-run them in a slightly
different way.

![Use arrow keys to go back to previous
commands.](//arrow_keys.png){width="100"}

## Challenge 4

What is the area of a circle with a radius of 5 cm?

Recall : ${circle = \pi r^2}$
`> 3.1416 * 5^2 [1] 78.54`\

R also has many built-in constants like pi.\
`> pi * 5^2 [1] 78.54` ++\