# Functions

## Introduction

Most of the time with R, you will need to use functions to do what you
want.

Functions are tools that are there to simplify your life. They allow you
to quickly execute operations on objects without having to write every
mathematical step. Functions are usually pre-existing R code that are
executed when used. They remove the need to create this code and to
rewrite it every time you need it.

To execute a function, you will need to call it. A function call
is essentially a shortcut to the code of the function.

To perform the function call you will need entry values called
**arguments** (or sometimes parameters). After performing its
operations, the function will then give you a **return value**. The
command also must be structured properly, following the "grammar
rules" of the R language (syntax).

A function call is structured as follow: the name of the function,
followed by parenthesis `( )`. Inside the parenthesis, you insert all
your arguments separated with commas.

`function_name(arg1, arg2, ...)`

Here is an example:

```{r, echo = TRUE, eval = FALSE}
sum(1, 2)
```

Arguments are **values** and the instructions the function needs to run.
Objects can be passed into functions:

```{r, echo = TRUE, eval = TRUE}
a <- 3
b <- 5
sum(a, b)
```

On the last line, the output that appears is the **return value** of the
function. In this case, it is the sum of `a` and `b`, 8.\

------------------------------------------------------------------------

### CHALLENGE 11

- Create a vector `a` that contains all numbers from 1 to 5

- Create an object `b` with value of 2

- Add `a` and `b` together using the basic `+` operator and save the
result in an object called `result_add`

- Add `a` and `b` together using the `sum` function and save the result in an object called `result_sum`

- Compare `result_add` and `result_sum`. Are they different?

- Add 5 to `result_sum` using the `sum` function.

#### Solution {-}

- Create a vector `a` that contains all numbers from 1 to 5

```{r, echo = TRUE, eval = TRUE}
a <- 1:5
```

- Create an object `b` with value of 2

```{r, echo = TRUE, eval = TRUE}
b <- 2
```

- Add `a` and `b` together using the basic `+` operator and save the
result in an object called `result_add`

```{r, echo = TRUE, eval = TRUE}
result_add <- a + b
```

- Add `a` and `b` together using the `sum` function and save the result in an object called `result_sum`

```{r, echo = TRUE, eval = TRUE}
result_sum <- sum(a,b)
```

- Compare `result_add` and `result_sum`. Are they different?

```{r, echo = TRUE, eval = TRUE}
result_add
```

The operation on the vector adds 2 to each element. The result is a
**vector**.

```{r, echo = TRUE, eval = TRUE}
result_sum
```

The function `sum()` adds all values of `a` and `b`. It is the same as
doing 1 + 2 + 3 + 4 + 5 + 2. The result is a **number**.

- Add 5 to `result_sum` using the `sum` function.

```{r, echo = TRUE, eval = TRUE}
sum(result_sum, 5)
```

------------------------------------------------------------------------

## Arguments

Arguments each have a **name** that can be provided during a function
call.\

To provide an argument name during a function call, just enter it in the
form `name=value`.

```{r, echo = TRUE, eval = FALSE}
log(8, base = 2)
```

If the name of the argument is not provided, the order of the arguments does
matter.\

```{r, echo = TRUE, eval = TRUE}
log(8, 2)
log(2, 8)
```

If the name argument is provided, the order of the arguments does not
matter.\

```{r, echo = TRUE, eval = TRUE}
log(x = 8, base = 2)
log(base = 2, x = 8)
```

------------------------------------------------------------------------

### CHALLENGE 12

`plot(x, y)` is a function that draws a graph of y as a function of x.
It requires two arguments named `x` and `y`.  
What are the differences between the following lines?

```{r, echo = TRUE, eval = FALSE}
a <- 1:100
b <- a^2
plot(a, b)
plot(b, a)
plot(x = a, y = b)
plot(y = b, x = a)
```

#### Solution {-}

```{r, echo = FALSE, eval = TRUE}
a <- 1:100
b <- a^2
```

```{r, echo = TRUE, eval = TRUE}
plot(a, b)
```

Plots the graph of `b` as a function of `a`.

```{r, echo = TRUE, eval = TRUE}
plot(b, a)
```

Plots the graph of `a` as a function of `b`.The argument names are not
provided, the order of the arguments matters!

```{r, echo = TRUE, eval = TRUE}
plot(x = a, y = b)
```

Plots the graph of `b` as a function of `a`, same as plot(a, b).

```{r, echo = TRUE, eval = TRUE}
plot(y = b, x = a)
```

Plots the graph of `b` as a function of `a`. The argument names are
provided, the order of the arguments does not matter!

------------------------------------------------------------------------

As a reference, here is a list of some of the most common R functions:

    sqrt, log, exp, max, min, sum, mean, sd, var, summary, plot, par, paste, format,
    head, length, str, names, typeof, class, attributes, library, ls, rm, setwd, getwd, 
    file.choose, c, seq, rep, tapply, lapply, aggregate, merge, cbind, rbind, unique,
    help (or ?), help.search (or ??), help.start